---
title: "RR Project 1"
output: html_document
---
#Reproducible Research: PA 1

##Load data:

```{r, echo=TRUE}
df<-read.csv("activity.csv")
print("num of NA values"); colSums(is.na(df))
```

##Total number of steps per day (ignoring NA values):
  * Calculate the total number of steps taken per day
  * Plot histogram of total steps per day
```{r, echo=TRUE}
sumsteps<-with(df,tapply(steps, date, FUN=sum))
print("sum of steps taken per day");sumsteps
hist(sumsteps, breaks = length(sumsteps)/4, main = "Distribution of Number of steps per day", 
    xlab = "Steps per day", ylab = "Frequency", col = "blue")
```

  * Calculate and report the mean and median of the total number of steps taken per day
```{r, echo=TRUE}
mean<-mean(sumsteps,na.rm=TRUE)
median<-median(sumsteps,na.rm=TRUE)
print("Mean number of steps per day");mean
print("Median number of steps per day");median
```
##What is the average daily activity pattern?
  * Time series plot of the 5-min interval and the average number of steps taken, averaged across all days (y-axis)
  * Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo=TRUE}
stepintervalmean<-with(df,tapply(steps,interval,FUN=mean,na.rm=TRUE))
plot(stepintervalmean,type='l',main = "Steps vs. Daily Average Interval",xlab = "Interval Index"
     ,ylab = "Number of Steps")
interv<-seq(along = stepintervalmean)[stepintervalmean == max(stepintervalmean)]
print("The Interval with the maximum number of steps is");interv
```
##Imputing missing values
  * Calculate and report the total number of missing values in the dataset.
  * Fill in all of the missing values in the dataset.
  * Create a new dataset that is equal to the original dataset but with the missing data filled in.
  * Make a histogram of the total number of steps taken each day and Calculate and report the mean      and median total number of steps taken per day. 
  * Do these values differ from the estimates from the first part of the assignment? What is the        impact of imputing missing data on the estimates of the total daily number of steps?
```{r, echo=TRUE}
nas<-sum(is.na(df$steps))
print("Number of missing values") ; nas 


# convert stepintervalmean to an array
simarray <- as.vector(stepintervalmean)
# Rep for each of the 61 days in order to achieve paralell index with df
simarray<- rep(simarray, 61)
# Set the non na values to 1
simarray[!is.na(df$steps)] <- 1

# Convert our df$steps into an array
steparray<-as.vector(df$steps)
# Set the missing values to 1
steparray[is.na(steparray)] <- 1

#multiply the arrays so that the imputed values will take the place of the na vals
# Add the imputed column to the original df
df$imputed <- simarray*steparray

impsteps <- with(df,tapply(imputed, date, FUN=sum))
hist(impsteps, breaks = length(impsteps/4), main = "Distribution of steps per day", 
    xlab = "Total steps per day", ylab = "Frequency", col = "red")
impmean<-mean(impsteps)
print("The mean of total steps for the imputed data is"); impmean
impmedian<-median(impsteps)
print("The mean of total steps for the imputed data is"); impmedian

#plot interval mean
impmeanint<-with(df,tapply(imputed,interval,mean))
plot(impmeanint, type = "l", xlab = "Interval index", ylab = "Numeber of Steps", 
    main = "Steps vs. Daily Average Interval")
```

##Are there differences in activity patterns between weekdays and weekends?

  * Create a new factor variable in the dataset with two levels - "weekday" and "weekend"         

  * Make a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.

```{r, echo=TRUE}
# Create a factor array where the variable is a factor of two levels weekdays, weekend-days
datefactor<-as.POSIXlt(df$date, format = "%Y-%m-%d")
days<- datefactor$wday
days[days == 0]<-0
days[days == 6]<-0
days[days != 0]<-1
factordays<-factor(days, levels = c(0, 1))
# Append factored array of days to the df
df$days<-factordays
# Calculate the mean
weekdaymean<-with(df,tapply(df$steps, list(interval,days), FUN=mean,na.rm=TRUE))

#create 2 side by side plots of the weekdays and weekends
par(mfrow = c(2, 1))
with(df, 
     {
    par(mai = c(0, 1, 1, 0))
    plot(weekdaymean[,1],type = "l",main=("Steps vs. Interval"),xaxt="n",ylab="Weekends")
    title = ("Number of Steps vs. Interval")
    par(mai = c(1, 1, 0, 0))
    plot(weekdaymean[, 2],type="l",xlab="Interval",ylab="Weekdays")

})

```






















